---
title: "Figure5_figure_replication"
author: "Michael Bowman and Bobby Bowman"
date: "2026-08-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Figure 5 replication
This script already assumes a seurat object and SupeRJump have been applied to for the dataset, we have these included in data("")


```{r load libraries}
library(SupeRJump)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(pals)


```

## load in completed seurat object and accompanying 

```{r load data}
load(system.file(paste0('data/barcoded_seurat_object.rDa'), package = 'SupeRJump'))


modified_Z<-Seurat::FetchData(seurat_obj, 
                              vars = c("custom_cell_classes_fine",
                                                 "sgRNA",
                                                 "custom_cell_classes_broad",
                                                 "sample","Combined_Ordering","absorbing_states"))



```



# Figure 5 B



```{r}

top_barcodes_to_use<-modified_Z%>%
  group_by(sgRNA)%>%
  mutate(barcode_count=n())%>%
  ungroup()%>%
  distinct(sgRNA,barcode_count)%>%
  arrange(desc(barcode_count))%>%
  arrange(desc(barcode_count))%>%
  #slice(2:11)%>% # the biggest "Barcode" is the no barcode group
  arrange(desc(barcode_count))%>%
  mutate(rank_order = row_number())%>%
  mutate(color_map=case_when(
     rank_order==2~"sgRNA_1",
     rank_order==3~"sgRNA_2",
     rank_order==4~"sgRNA_3",
     rank_order==5~"sgRNA_4",
     rank_order==6~"sgRNA_5",
     rank_order==7~"sgRNA_6",
     rank_order==8~"sgRNA_7",
     rank_order==9~"sgRNA_8",
     rank_order==10~"sgRNA_9",
     rank_order==11~"sgRNA_10",
     TRUE~"others"))%>%
  mutate(color_map=factor(color_map,levels=c(paste0("sgRNA_",1:10),"others")))

ggplot(top_barcodes_to_use%>%
         filter(rank_order!=1)%>%arrange(desc(color_map)),
       aes(x=rank_order,y=barcode_count,color=color_map))+
  geom_point(size=3)+
  scale_color_manual(values = c("others"="grey70",
     "sgRNA_1"=pals::cols25(n=11)[2],
     "sgRNA_2"=pals::cols25(n=11)[3],
     "sgRNA_3"=pals::cols25(n=11)[4],
    "sgRNA_4"=pals::cols25(n=11)[5],
    "sgRNA_5"=pals::cols25(n=11)[1],
     "sgRNA_6"=pals::cols25(n=11)[7],
    "sgRNA_7"=pals::cols25(n=11)[8],
    "sgRNA_8"=pals::cols25(n=11)[9],
     "sgRNA_9"=pals::cols25(n=24)[24],
     "sgRNA_10"=pals::cols25(n=16)[16]))+
  geom_hline(yintercept = 5,linetype = "dashed",linewidth=1.4)+
  ggpubr::theme_classic2(base_size = 18)+
  theme(legend.title = element_blank())+
  ylim(c(0,600))+
  xlab("Abundance Ranking")+
  ylab("Cell abundance per barcode")


```


# Figure 5C 
 This produces a stacked barplot 


```{r}
top_10_barcodes_to_use<-modified_Z%>%
  group_by(sgRNA)%>%
  mutate(barcode_count=n())%>%
  ungroup()%>%
  distinct(sgRNA,barcode_count)%>%
  arrange(desc(barcode_count))%>%
  arrange(desc(barcode_count))%>%
  slice(2:11)%>% # the biggest "Barcode" is the no barcode group
  arrange(desc(barcode_count))%>%
  pull(sgRNA)


table_values<-modified_Z%>%
  filter(sgRNA%in%top_10_barcodes_to_use)%>%
 dplyr::select(sgRNA,custom_cell_classes_broad)

table_values<-table(table_values$sgRNA,table_values$custom_cell_classes_broad)
table_values<-table_values/rowSums(table_values)

  data.frame(table_values[top_10_barcodes_to_use,])%>%
    dplyr::rename(sgRNA="Var1",
                  Cell="Var2",
                  Proportion="Freq")%>%
  mutate(sgRNA=factor(sgRNA,levels=top_10_barcodes_to_use))%>%
  mutate(sgRNA = paste0("sgRNA_",as.integer(sgRNA)-1))%>%
    mutate(sgRNA=ifelse(sgRNA=="sgRNA_0","Unk.",sgRNA))%>%
  mutate(sgRNA = factor(sgRNA,levels=unique(sgRNA)))%>%
  mutate(Cell=factor(Cell,levels=c("HSPC","MultiLin","MDP","BMCP","Erp","MkP","Baso","Mast","Neu","MP","cDC","Lymph")))%>%
  ggplot(aes(x=sgRNA,y=Proportion,fill=Cell))+
  geom_col(position="stack",show.legend = T)+
  ggpubr::theme_classic2(base_size=24)+
  scale_y_continuous(expand=c(0,0),limits = c(0, NA))+
    scale_fill_manual(values=pals::watlington())+
    guides(fill = guide_legend(title = ""))+
theme(axis.title.x = element_blank(),
      axis.text.x=element_text(angle=45,hjust=1,vjust=1,size=16))

```

# Figure 5 D
We show visitation probability going from from finer cell types (rows) to coarser ones (columns). 
visitation probability should not incldue absorbing states, so be sure to eliminate those cells from both the rows and columns.

```{r figure 5D visitation probability}

combined_data_VP<-rbind(reshape2::melt(as.matrix(seurat_obj@graphs$visitation),
                                       varnames=c("current_state","next_state"),value.name="probability"))%>%
  inner_join(modified_Z%>%
               filter(is.na(absorbing_states))%>%
               tibble::rownames_to_column(var="current_state")%>%
               select(current_state,custom_cell_classes_broad,sgRNA,custom_cell_classes_fine,sample))%>%
  rename(cluster_init=custom_cell_classes_broad)%>%
  rename(init_barcode=sgRNA)%>%
  rename(init_fine_cells=custom_cell_classes_fine)


combined_data_VP<-combined_data_VP%>%
  inner_join(modified_Z%>%
               filter(is.na(absorbing_states))%>%
               tibble::rownames_to_column(var="next_state")%>%
               select(next_state,custom_cell_classes_broad,sgRNA,custom_cell_classes_fine))%>%
  rename(cluster_to=custom_cell_classes_broad)%>%
 rename(end_barcode=sgRNA)%>%
  rename(end_fine_cells=custom_cell_classes_fine)

combined_data_VP<-combined_data_VP%>%
  dplyr::group_by(cluster_init,cluster_to)%>%
  dplyr::mutate(red_mean=mean(probability))%>%
  dplyr::ungroup()%>%
  dplyr::group_by(cluster_init,cluster_to,init_barcode)%>%
  dplyr::mutate(full_mean_sgRNA=mean(probability))%>%
  dplyr::ungroup()%>%
  dplyr::mutate(fitted_lm_mean_adjusted_sgRNA=probability-(full_mean_sgRNA-red_mean))%>%
  mutate(scaled_probability_sgRNA=(fitted_lm_mean_adjusted_sgRNA-min(fitted_lm_mean_adjusted_sgRNA))/(max(fitted_lm_mean_adjusted_sgRNA)-min(fitted_lm_mean_adjusted_sgRNA)))



temp_prob_blocks_sgRNA5<-combined_data_VP%>%
  group_by(init_fine_cells,cluster_to)%>%
  summarize(block_mean=sum(scaled_probability_sgRNA))%>%
  mutate(init_fine_cells=factor(init_fine_cells,levels=c("MPP1","MPP2","MPP3","MPP4","MPP5","ML1","ML2","MEP","early_Erp","MkP","MDP","MP","cDC","BMCP","Mast","Baso","Neu","Lymph")))%>%
  dplyr::mutate(cluster_to = factor(cluster_to,levels=c("HSPC","MultiLin","MDP","BMCP","Erp","MkP","Baso","Mast","Neu","MP","cDC","Lymph")))%>%
  arrange(init_fine_cells,cluster_to)%>%
  pivot_wider(id_cols = init_fine_cells,names_from=cluster_to,values_from=block_mean)%>%
  tibble::column_to_rownames(var="init_fine_cells")

Cluster_to_cluster_jump_VP_sgRNA5<-apply(temp_prob_blocks_sgRNA5,2,function(x){x/sum(x)})
Cluster_to_cluster_jump_VP_sgRNA5<-t(apply(Cluster_to_cluster_jump_VP_sgRNA5,1,function(x){scale(x)}))
colnames(Cluster_to_cluster_jump_VP_sgRNA5)<-colnames(temp_prob_blocks_sgRNA5)

pheatmap::pheatmap(Cluster_to_cluster_jump_VP_sgRNA5,
                   color=grDevices::colorRampPalette(colors = rev(RColorBrewer::brewer.pal(n = 11, name = "RdBu")))(100),
                   scale="none",
                   cluster_rows = F,
                   cluster_cols = F,
                   border_color = "white",
                    cellwidth = 20,
                   cellheight = 20,filename = "./Figures/VisitationProbability_fine_to_broad_for_all_cells.pdf")

```



# Figure 5 E

```{r}

combined_data<-rbind(reshape2::melt(t(GetAssayData(seurat_obj,assay = "lineage_fates"))%>%as.matrix(),
                                    varnames=c("current_state","cluster_to"),
                                    value.name="probability"))%>%
  inner_join(modified_Z%>%
               tibble::rownames_to_column(var="current_state")%>%
               select(current_state,custom_cell_classes_broad,sgRNA,custom_cell_classes_fine,sample))

Matrix_heatmap_data<-combined_data%>%
  mutate(cluster_to=factor(cluster_to,levels=c("Erp","MkP","Baso","Mast","Neu","MP","cDC","Lymph")))%>%
  filter(custom_cell_classes_broad=="HSPC"|custom_cell_classes_broad=="MultiLin")%>%
  filter(sgRNA %in%top_10_barcodes_to_use)%>%
  group_by(custom_cell_classes_broad,sgRNA,cluster_to)%>%
  summarize(med_value_raw = median(probability))%>%
  ungroup()%>%
  mutate(sgRNA = factor(sgRNA,levels=top_10_barcodes_to_use))%>%
  arrange(sgRNA)%>%
  mutate(sgRNA = paste0("sgRNA_",as.integer(sgRNA)))%>%
  mutate(sgRNA = factor(sgRNA,levels=unique(sgRNA)))%>%  
  mutate(column_index_for_heatmap = paste0(custom_cell_classes_broad,"_",sgRNA))%>%
  mutate(column_index_for_heatmap= factor(column_index_for_heatmap,
                                          levels=c(paste0("HSPC_sgRNA_",0:10),
                                                   paste0("MultiLin_sgRNA_",0:10))))%>%
  arrange(column_index_for_heatmap)

list_of_colors_annotation <- list(
  `cell type`=c(`MultiLin`=pals::watlington(n=12)[2],
          `HSPC`=pals::watlington(n=12)[1]),
  cluster_to = c(`Erp`=pals::watlington(n=12)[5],
            `MkP`=pals::watlington(n=12)[6],
            `Baso`=pals::watlington(n=12)[7],
            `Mast`=pals::watlington(n=12)[8],
            `Neu`=pals::watlington(n=12)[9],
            `MP`=pals::watlington(n=12)[10],
            `cDC`=pals::watlington(n=12)[11],
            `Lymph`=pals::watlington(n=12)[12]),
  guide = c(
              `sgRNA_1`=pals::cols25(n=11)[2],
              `sgRNA_2`=pals::cols25(n=11)[3],
              `sgRNA_3`=pals::cols25(n=11)[4],
              `sgRNA_4`=pals::cols25(n=11)[5],
              `sgRNA_5`=pals::cols25(n=11)[1],
              `sgRNA_6`=pals::cols25(n=11)[7],
           `sgRNA_7`=pals::cols25(n=11)[8],
           `sgRNA_8`=pals::cols25(n=11)[9],
           `sgRNA_9`=pals::cols25(n=24)[24],
           `sgRNA_10`=pals::cols25(n=16)[16])
)

tidyheatmaps::tidy_heatmap(df = Matrix_heatmap_data%>%as.data.frame%>%
                             filter(sgRNA !="sgRNA_0")%>%
                             dplyr::rename(guide="sgRNA")%>%
                             dplyr::rename(`cell type`="custom_cell_classes_broad"),
                           rows = cluster_to,
                           columns = column_index_for_heatmap,
                          values = med_value_raw,
                           annotation_col = c(`cell type`,guide),
                           annotation_colors = list_of_colors_annotation,
                           border_color = "white",
                           colors = colorRampPalette(rev(RColorBrewer::brewer.pal(11, "RdBu")))(100),
                           scale = "row",display_numbers = F,
                           color_legend_n = 100,gaps_col = `cell type`,
                           fontsize = 12,cellwidth = 12,cellheight = 12,
                           show_rownames = T,show_colnames = F,
                           angle_col = 45,legend = T,annotation_legend = T,
                           cluster_rows = F,cluster_cols = F,annotation_names_col = T,filename = "./Figures/Fate_heatmap_with_legend.pdf",width=6,height=9)




anova_model <- aov(mean_fate ~ custom_cell_classes_broad * alternative_boxplot_cats, data = check_fates_df_truncated%>%
         distinct(cell_names,alternative_boxplot_cats,mean_fate,.keep_all = T)%>%
         mutate(mean_fate = (mean_fate-min(mean_fate))/(max(mean_fate)-min(mean_fate))))
summary(anova_model)

ANOVA_TUKEY_P_VAL<-stats::TukeyHSD(anova_model, which = "custom_cell_classes_broad:alternative_boxplot_cats")

```



# Figure 5 F
This figure looks into inter, intra, vs alternative routes to a cDC lineage. 

```{r}


check_fates_df<-seurat_obj@misc$fates_individual%>%
  data.frame%>%
  tibble::rownames_to_column(var="cell_names")%>%
  pivot_longer(cols = !c(cell_names),names_to = "sink_cell",values_to = "Fate")%>%
  filter(Fate!=0)

check_fates_df<-check_fates_df%>%
  inner_join(modified_Z%>%
               filter(is.na(absorbing_states))%>%
               tibble::rownames_to_column(var="cell_names"),
             by="cell_names")

check_fates_df<-check_fates_df%>%
  inner_join(modified_Z%>%
              filter(!is.na(absorbing_states))%>%
               tibble::rownames_to_column(var="cell_names")%>%
               dplyr::rename(sink_cell="cell_names"), by="sink_cell",suffix=c("","_sink_cells"))



check_fates_df_truncated<-check_fates_df%>%
  mutate(sgRNA=ifelse(is.na(sgRNA),"none",sgRNA))%>%
  mutate(sgRNA_sink_cells=ifelse(is.na(sgRNA_sink_cells),"none",sgRNA_sink_cells))%>%
  filter(custom_cell_classes_fine_sink_cells%in%c("cDC"))%>%
  filter(sgRNA!="none")%>%
  filter(custom_cell_classes_broad%in%c("MultiLin","HSPC","cDC"))%>%
  mutate(custom_cell_classes_broad=factor(custom_cell_classes_broad,levels=c("HSPC","MultiLin","cDC")))%>%
  mutate(alternative_boxplot_cats = case_when(sgRNA==sgRNA_sink_cells~"Intra",
                                              (sgRNA%in%c(unique(sgRNA_sink_cells)))&(sgRNA!=sgRNA_sink_cells)~"Inter",                                                      !(sgRNA%in%c(unique(sgRNA_sink_cells)))&(sgRNA!=sgRNA_sink_cells)~"Alt.",
                                              TRUE~"SKIP"))%>%
  mutate(alternative_boxplot_cats=factor(alternative_boxplot_cats,levels=c("Alt.","Intra","Inter")))%>%
  filter(Fate!=1 & Fate!=0)%>%
  group_by(cell_names,alternative_boxplot_cats)%>%
  mutate(mean_fate = mean(Fate))%>%
  ungroup()

ggplot(check_fates_df_truncated%>%
         distinct(cell_names,alternative_boxplot_cats,mean_fate,.keep_all = T)%>%
         mutate(mean_fate = (mean_fate-min(mean_fate))/(max(mean_fate)-min(mean_fate)))%>%
         mutate(custom_cell_classes_broad=factor(custom_cell_classes_broad)),
       aes(x=alternative_boxplot_cats,y=mean_fate,fill=alternative_boxplot_cats))+
  geom_boxplot(show.legend = F)+
  facet_wrap(~custom_cell_classes_broad,ncol = 4)+
  ggpubr::theme_classic2(base_size = 14)+
  theme(axis.text.x = element_text(angle=45,hjust=1,vjust=1), 
        axis.title.x = element_blank(), 
        strip.background = element_blank(),)+
  scale_fill_manual(values=pals::okabe(n=3))+ 
  ylab("Normalized\n Fate Score")+
  scale_x_discrete(labels= c("Alt.","Intra","Inter"))


ggsave(filename = "./Figures/fate_boxplots_guide_to_guides.pdf",plot = last_plot(),width = 4.875,height=2.75)



anova_model <- aov(mean_fate ~ custom_cell_classes_broad * alternative_boxplot_cats, data = check_fates_df_truncated%>%
         distinct(cell_names,alternative_boxplot_cats,mean_fate,.keep_all = T)%>%
         mutate(mean_fate = (mean_fate-min(mean_fate))/(max(mean_fate)-min(mean_fate))))
summary(anova_model)


```


# Figure 5 G

This figure looks into CMFPT vs the Visitation probability for cells going to the highest fated MultiLin


```{r}

cell_id<-check_fates_df_truncated%>%
         select(cell_names,custom_cell_classes_broad,Fate,alternative_boxplot_cats,mean_fate)%>%
         distinct(cell_names,alternative_boxplot_cats,Fate,.keep_all = T)%>%
  filter(custom_cell_classes_broad=="MultiLin")%>%
  filter(alternative_boxplot_cats=="Intra")%>%
  arrange(desc(Fate))%>%
  slice(1)%>%
  pull(cell_names)

temp_prob_vec_go<-combined_data_VP%>%filter(next_state==cell_id)

temp_time_vec_go<-reshape2::melt(as.matrix(seurat_obj@graphs$CMFPT),varnames=c("current_state","next_state"),value.name="CMFPT")%>%
  
  inner_join(modified_Z%>%
               tibble::rownames_to_column(var="current_state")%>%
               select(current_state,custom_cell_classes_broad,sgRNA))%>%
  rename(cluster_init=custom_cell_classes_broad)%>%
  
  inner_join(modified_Z%>%
               tibble::rownames_to_column(var="next_state")%>%
               select(next_state,custom_cell_classes_broad))%>%
  rename(cluster_to=custom_cell_classes_broad)%>%
  
  filter(next_state==cell_id)

combined_df_to_plot<-inner_join(temp_prob_vec_go,
                                temp_time_vec_go,
                                by=c("current_state","next_state","cluster_init","cluster_to"))%>%
  mutate(color_class=case_when(init_barcode=="TCAATCTATCTAACTTGGAA"~"sgRNA_5",
                               init_barcode=="TCTTTGGTAGGTAGCTGCTT"~"sgRNA_4",
                                init_barcode=="CGTAGGTTCGTTCCTATCGT"~"sgRNA_9",
                               init_barcode=="CGGTACCAGCATAGGTTGCA"~"sgRNA_10",
                               TRUE~"Other"))%>%
  mutate(color_class=factor(color_class,levels=c("Other","sgRNA_4","sgRNA_5","sgRNA_9","sgRNA_10")))%>%
  arrange(color_class)%>%
  filter(CMFPT!=0)%>%
  filter(cluster_init%in%c("MultiLin","HSPC"))%>%
  filter(init_barcode!="none")%>%
  mutate(cluster_init= factor(cluster_init,levels=c("HSPC","MultiLin")))

ggplot(combined_df_to_plot,aes(x=log10(CMFPT+1),y= probability,color=color_class))+
  geom_point(show.legend = T,alpha=0.7)+

  ggpubr::theme_classic2(base_size = 20)+
  scale_color_manual(values = c("Other"="grey70",
                                "sgRNA_5"=pals::cols25(n=6)[1],
                                "sgRNA_9"=pals::cols25(n=24)[24],
                                "sgRNA_4"=pals::cols25(n=6)[5],
                                "sgRNA_10"=pals::cols25(n=16)[16]))+
  theme(legend.title = element_blank(),
        legend.position = "top",
          strip.background = element_blank())+
  ylab("Visitation Score")+
  xlab(expression(log[10](CMFPT)))+
  facet_wrap(~cluster_init,axes="all",axis.labels = "margins")


ggsave(filename = "./Figures/CMFPT_vs_VP_from_barcodes_to_highest_fated_multilin_cell_split_by_broad_with_guide4_5_9_10.pdf",width = 8,height=5)


summary(manova(cbind(log10(CMFPT+1),scaled_probability_sample)~color_class:cluster_init,combined_df_to_plot))

library(RVAideMemoire)
pw <- pairwise.perm.manova(  cbind(log10(combined_df_to_plot$CMFPT+1), combined_df_to_plot$scaled_probability_sample),interaction(combined_df_to_plot$color_class,combined_df_to_plot$cluster_init),  nperm = 999)



```



# Figure 5 H

```{r}


combined_data_flight_time_wide_matrix<-reshape2::melt(as.matrix(seurat_obj@graphs$CMFPT),varnames=c("current_state","next_state"),value.name="CMFPT")%>%
  
  inner_join(modified_Z%>%
               tibble::rownames_to_column(var="current_state")%>%
               select(current_state,custom_cell_classes_broad,sgRNA))%>%
  rename(cluster_init=custom_cell_classes_broad)%>%
  
  inner_join(modified_Z%>%
               tibble::rownames_to_column(var="next_state")%>%
               select(next_state,custom_cell_classes_broad))%>%
  rename(cluster_to=custom_cell_classes_broad)%>%
  
  dplyr::group_by(cluster_init,cluster_to)%>%
  dplyr::ungroup()%>%
  dplyr::group_by(cluster_init,cluster_to,sgRNA)%>%
  dplyr::ungroup()%>%
  mutate(log_CMFPT=log10(CMFPT+1))%>%
  dplyr::select(current_state,next_state,log_CMFPT)%>%
  tidyr::pivot_wider(id_cols = current_state,names_from = next_state,values_from = log_CMFPT,values_fill = Inf)%>%
  tibble::column_to_rownames(var="current_state")


combined_data_VP_wide_matrix<-combined_data_VP%>%
  dplyr::select(current_state,next_state,scaled_probability_sgRNA)%>%
  tidyr::pivot_wider(id_cols = current_state,names_from = next_state,values_from = scaled_probability_sgRNA,values_fill = 0)%>%
  tibble::column_to_rownames(var="current_state")

# A quick trick to update the visitation graph with our custom normalization outside of regular batch correction for WDT enrichment
G <- seurat_obj@graphs[["visitation"]]
diag(combined_data_VP_wide_matrix)<-0
dense_cells <- rownames(combined_data_VP_wide_matrix)
idx <- match(dense_cells, rownames(G))
G[idx, idx] <- as.matrix(combined_data_VP_wide_matrix)

seurat_obj@graphs[["visitation"]] <- G
 
seurat_obj<-SupeRJump::GetWeightedDestinationTime(seurat_obj = seurat_obj,
                                       preference_lineage_column = "sgRNA",
                                       enrichment_flag = T)


E_df<-seurat_obj@misc[["WDT_pref_enrichment"]][top_10_barcodes_to_use,]%>%
  as.data.frame%>%
  tibble::rownames_to_column(var="guide")%>%
  mutate(guide = factor(guide,levels=top_10_barcodes_to_use))%>%
  mutate(guide = paste0("sgRNA_",as.integer(guide)))%>%
  mutate(guide = factor(guide,levels=unique(guide)))%>%  
  mutate(guide=factor(guide,levels=rev(c("sgRNA_9","sgRNA_7","sgRNA_4","sgRNA_2","sgRNA_1","sgRNA_3","sgRNA_6","sgRNA_8","sgRNA_5","sgRNA_10"))))%>%
  pivot_longer(cols = !c(guide),names_to = "lineage",values_to="WDT_enrich")%>%
  mutate(lineage = gsub("_membership","",lineage))%>%
  group_by(lineage)%>%
  arrange(desc(WDT_enrich))%>%
  mutate(guide_rank = row_number())%>%
  ungroup()%>%
  arrange(guide)%>%
  mutate(lineage = factor(lineage,levels=c("HSPC","MultiLin","MDP","BMCP","Erp","MkP","Baso","Mast","Neu","MP","cDC","Lymph")))
  
  

 ggplot(E_df%>%
          filter(lineage=="Erp"|lineage=="Lymph")%>%
          arrange(lineage,guide_rank)%>%
          arrange(guide),
        aes(y=guide,x=guide_rank,color=lineage))+
   geom_point(size=3)+
   ggpubr::theme_classic2(base_size=18)+
   theme(axis.title.y = element_blank(),
         legend.title = element_blank())+
   xlab("WDT enrichment rank")+
   scale_color_manual(values=c("Erp"=pals::watlington(n=12)[5],
                               "Lymph"=pals::watlington(n=12)[12]))
 
 
 cor.test(E_df%>%
          filter(lineage=="Erp")%>%
          arrange(lineage,guide)%>%pull(WDT_enrich),
 E_df%>%filter(lineage=="Lymph")%>%
   arrange(lineage,guide)%>%
   pull(WDT_enrich),
 alternative = "two.sided",method="spearman")

```


